An Embedding Framework for the Design and Analysis of Consistent Polyhedral Surrogates

### 2022-05-31 poly-embed-journal/poly-embed-journal.tex
section 6.0-6.4 pass

We have shown in \S~\ref{sec:poly-loss-embed} a tight connection between embeddings and polyhedral losses.
Here we go beyond polyhedral losses, showing a more general necessary condition for an embedding: a surrogate embeds a discrete loss if and only if it has a polyhedral Bayes risk, or equivalently, a finite representative set (Lemma~\ref{lem:X}).
This result implies that the embedding condition simplifies to matching Bayes risks (Proposition~\ref{prop:embed-bayes-risks}).
We also use these result to understand deeper structure of embeddings, even down to the geometry of the underlying properties. %, and the equivalence of various notions of non-redundant predictions.
In particular, we study a natural notion of a ``trimmed'' loss function (Definition~\ref{def:trim-loss}), and connect this definition to both tight embeddings and non-redundancy from property elicitation (Proposition~\ref{prop:embed-iff-trims-equal}).

@@

%% generalizing from structure of polyhedral surrogates to surrogates with polyhedral risk (and why not necessarily the same)
While we have focused on polyhedral losses thus far, many of our results about embeddings extend to losses with polyhedral Bayes risks, a weaker condition.
(We say a concave function is polyhedral if its negation is a polyhedral convex function.)
To see that every polyhedral loss has a polyhedral Bayes risk, recall that Theorem~\ref{thm:poly-embeds-discrete} constructs a finite representative set $\Sc$ for any polyhedral loss $L$, and thus $\risk{L} = \risk{L|_\Sc}$ by Lemma~\ref{lem:loss-restrict}, which is polyhedral.
The condition is strictly weaker: a Bayes risk may be polyhedral even if the loss itself is not.
For example, a modified hinge loss $L(r)_y = \max(r^2-1,1-ry)$
as shown in Figure~\ref{fig:modified-hinge}, which matches hinge loss on the interval $[-1,1]$ but is strictly convex outside the interval $[-2,2]$, still embeds twice 0-1 loss.

\begin{figure}
	\begin{minipage}{0.47\linewidth}
		\centering
		\includegraphics[width=0.95\linewidth]{figs/modhinge.pdf}
		%		\caption{Bayes risk of the 0-1 loss.}
		%		\label{fig:0-1-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\linewidth}
		\centering		\includegraphics[width=0.95\linewidth]{figs/modhinge-br.pdf}
		%		\caption{Bayes risk of hinge loss.}
		%		\label{fig:hinge-br}
	\end{minipage}
	\caption{(L) Expected modified hinge loss for fixed distribution; (R) Bayes risk of modified hinge still matches the Bayes risk of hinge.}
	\label{fig:modified-hinge}
\end{figure}

% \begin{lemma}
% 	\label{lem:poly-loss-poly-risk}
% 	If $L$ is polyhedral, $\risk{L}$ is polyhedral.
% \end{lemma}
% \begin{proof}
% 	Let $L:\reals^d\to\reals_+^\Y$ be a polyhedral loss, and $\Gamma = \prop{L}$.
% 	By Lemma~\ref{lem:polyhedral-range-gamma}, $\U = \Gamma(\simplex)$ is finite. 
% 	For each $U\in \U$, select $u_U \in U$, and let $\Sc = \{u_U : U \in\U\}$.
% 	$\Sc$ is representative for $L$, so Lemma~\ref{lem:loss-restrict} gives us $\risk{L} = \risk{L|_{\Sc}}$, which is polyhedral as $\Sc$ is finite.
% \end{proof}

%% intro to Lemma X
We now present our main structural result in Lemma~\ref{lem:X}, which we prove in \S~\ref{app:polyhedra}.
%This result lays the foundation for the rest of this section.
Lemma~\ref{lem:X} observes that (minimizable) losses $L$ with polyhedral Bayes risk have finite representative sets, and derives equivalent conditions on the level sets of the $\prop{L}$ and discrete losses tightly embedded by $L$.
\raft{TODO: Brief description of the result}\jessiet{Attempted. burn if okay}

@@

Lemma~\ref{lem:X} leads to another appealing equivalent condition to our embedding condition in Definition~\ref{def:loss-embed}: a surrogate embeds a discrete loss if and only if their Bayes risks match.

@@

  $\implies$
  Suppose $L$ embeds $\ell$, so we have some $\Sc\subseteq \R$ which is representative for $\ell$ and an embedding $\varphi:\Sc\to\reals^d$; take $\U := \varphi(\Sc)$.
  Since $\Sc$ is representative for $\ell$, by embedding condition (ii) we have $\{\gamma_s \mid s\in\Sc\} = \{\Gamma_u \mid u\in\U\}$, so $\U$ is representative for $L$.
  % for all $p\in\simplex$ we have some $r \in \Sc \cap \gamma(p)$ and thus we have $\varphi(r) \in \U \cap \Gamma(p)$.
  % In particular, $\U \cap \Gamma(p)\neq \emptyset$ for all $p\in\simplex$, 
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{\ell} = \risk{\ell|_{\Sc}}$ and $\risk{L} = \risk{L|_{\U}}$.
  As $L(\varphi(\cdot)) = \ell(\cdot)$ by embedding condition (i), for all $p\in\simplex$ we have
  \begin{equation*}
    \risk{\ell}(p) = \risk{\ell|_\Sc}(p) = \min_{r \in \Sc}\inprod{p}{\ell(r)} = \min_{r \in \Sc}\inprod{p}{L(\varphi(r))} = \min_{u \in \U}\inprod{p}{L(u)} = \risk{L|_\U}(p) = \risk{L}(p)~.
  \end{equation*}

  $\impliedby$
	For the reverse implication, assume $\risk{L} = \risk{\ell}$, which are polyhedral functions as $\ell$ is discrete.
  From Lemma~\ref{lem:X}(\ref{item:X-min-V}), we have some set $\V\subseteq\reals^\Y_+$ and minimum representative sets $\R^* \subseteq \R$ and $\U^* \subseteq \U$, for $\ell$ and $L$ respectively, such that $\ell(\R^*) = \V = L(\U^*)$.
  %\jessiet{This is a different $\Sc$ than the first paragraph, right? The first $\Sc$ is representative for $\ell$.}\raft{They are both quantified separately, so not linked; are you saying it's confusing to use $\Sc$ here though because its role is slightly different?}\jessiet{Yes... attempting to clarify.}
  As $\R^*$ and $\U^*$ are miniumum, they cannot repeat loss vectors, and thus $|\R^*|=|\ell(\R^*)|$ and $|L(\U^*)|=|\U^*|$.
  We conclude that $\R^*$ and $\U^*$ are both in bijection with $\V$.
  The map $\varphi :\R^* \to \reals^d$, given by $\varphi(r) = u \in \U^*$ where $\ell(r) = L(u)$, is therefore well-defined.
  Condition (i) of an embedding is immediate.
  From Proposition~\ref{prop:representative-embeds-restriction}, $\ell$ embeds $\ell|_{\R^*}$ and $L$ embeds $L|_{\U^*}$, both via the identity embedding.
  Using condition (ii) from both embeddings, for all $p\in\simplex$ and $r\in\R^*$, we have
  \begin{equation*}
    r \in \gamma(p) \iff r \in \prop{\ell|_{\R^*}}(p) \iff \varphi(r) \in \prop{L|_{\U^*}}(p)
    \iff \varphi(r) \in \prop{L}(p)~,
  \end{equation*}
  giving condition (ii).
\end{proof}

@@

\raf{Short parag on why we care about trim}\jessie{Attempted; burn? i don't like this.}
Trimmed losses allow us to understand ``where the action happens'' even in the best case scenario where the loss vector is optimal for \emph{some} data distribution. \jessie{phrasing}
In what follows, we relate the trimmed form of a loss to the concept of redundancy in property elicitation.

@@

\section{Polyhedral Indirect Elicitation Implies Consistency\jessie{Calibration?}}
\label{sec:poly-ie-consistency}

>>>>

We have shown in \S~\ref{sec:poly-loss-embed} a close connection between embeddings and polyhedral losses.
Here we go beyond polyhedral losses, showing a more general necessary condition for an embedding: a surrogate embeds a discrete loss if and only if it has a polyhedral Bayes risk, or equivalently, a finite representative set (Lemma~\ref{lem:X}).
This result implies that the embedding condition simplifies to matching Bayes risks (Proposition~\ref{prop:embed-bayes-risks}).
We also use this result to understand deeper structure of embeddings, and the geometry of the underlying properties. %, and the equivalence of various notions of non-redundant predictions.
In particular, we study a natural notion of a ``trimmed'' loss function (Definition~\ref{def:trim-loss}), and connect this notion to tight embeddings, and to non-redundancy from property elicitation (Proposition~\ref{prop:embed-iff-trims-equal}).

@@

%% generalizing from structure of polyhedral surrogates to surrogates with polyhedral risk (and why not necessarily the same)
While we have focused on polyhedral losses thus far, many of our results extend to losses with polyhedral Bayes risks, a strictly weaker condition.
(We say a concave function is polyhedral if its negation is a polyhedral convex function.)
To see that every polyhedral loss has a polyhedral Bayes risk, recall that Theorem~\ref{thm:poly-embeds-discrete} constructs a finite representative set $\Sc$ for any polyhedral loss $L$, and thus $\risk{L} = \risk{L|_\Sc}$ by Lemma~\ref{lem:loss-restrict}, which is polyhedral.
Conversely, however, a Bayes risk may be polyhedral even if the loss itself is not.
For example, a modified hinge loss $L(r)_y = \max(r^2-1,1-ry)$
as shown in Figure~\ref{fig:modified-hinge}, which matches hinge loss on the interval $[-1,1]$ but is strictly convex outside the interval $[-2,2]$, still embeds twice 0-1 loss.

%% intro to Lemma X
Much of our embedding framework relies on the existence of finite representative sets.
Our main structural result is that a minimizable loss has a finite representative sets if and only if its Bayes risk is polyhedral.
The proof looks at the facets (full-dimensional faces) of the Bayes risk, and argues that each facet is generated by the loss at a particular report, and the (finite) set of these reports is representative.
Along the way, we identify several other useful facts deriving from this same geometry; for example, a discrete loss tightly embedded by a loss are unique up to relabeling, any set-wise minimal representative set must be minimum in cardinality, and the level sets of the corresponding property are unique and full-dimensional.
Together, these facts form Lemma~\ref{lem:X}, which we use throughout this section.
See \S~\ref{app:polyhedra} for ommitted proofs.

\begin{figure}[t]
	\begin{minipage}{0.47\linewidth}
		\centering
		\includegraphics[width=0.95\linewidth]{figs/modhinge.pdf}
		%		\caption{Bayes risk of the 0-1 loss.}
		%		\label{fig:0-1-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\linewidth}
		\centering		\includegraphics[width=0.95\linewidth]{figs/modhinge-br.pdf}
		%		\caption{Bayes risk of hinge loss.}
		%		\label{fig:hinge-br}
	\end{minipage}
	\caption{(L) Expected modified hinge loss for fixed distribution; (R) Bayes risk of modified hinge still matches the Bayes risk of hinge.}
	\label{fig:modified-hinge}
\end{figure}

@@

Lemma~\ref{lem:X} leads to another appealing equivalent condition to an embedding: a surrogate embeds a discrete loss if and only if their Bayes risks match.
The proof follows by mapping the two conditions of an embedding onto the geometric structure revealed by Lemma~\ref{lem:X}: (ii) if the properties have the same level sets, the Bayes risks have the same projections onto $\simplex$, and (i) if the loss values match, then the slopes of the Bayes risk must be identical as well.

@@

%  $\implies$
  Suppose $L$ embeds $\ell$, so we have some $\Sc\subseteq \R$ which is representative for $\ell$ and an embedding $\varphi:\Sc\to\reals^d$; take $\U := \varphi(\Sc)$.
  Since $\Sc$ is representative for $\ell$, by embedding condition (ii) we have $\{\gamma_s \mid s\in\Sc\} = \{\Gamma_u \mid u\in\U\}$, so $\U$ is representative for $L$.
  % for all $p\in\simplex$ we have some $r \in \Sc \cap \gamma(p)$ and thus we have $\varphi(r) \in \U \cap \Gamma(p)$.
  % In particular, $\U \cap \Gamma(p)\neq \emptyset$ for all $p\in\simplex$, 
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{\ell} = \risk{\ell|_{\Sc}}$ and $\risk{L} = \risk{L|_{\U}}$.
  As $L(\varphi(\cdot)) = \ell(\cdot)$ by embedding condition (i), for all $p\in\simplex$ we have
  \begin{equation*}
    \risk{\ell}(p) = \risk{\ell|_\Sc}(p) = \min_{r \in \Sc}\inprod{p}{\ell(r)} = \min_{r \in \Sc}\inprod{p}{L(\varphi(r))} = \min_{u \in \U}\inprod{p}{L(u)} = \risk{L|_\U}(p) = \risk{L}(p)~.
  \end{equation*}

%  $\impliedby$
	For the reverse implication, assume $\risk{L} = \risk{\ell}$, which are polyhedral functions as $\ell$ is discrete.
  From Lemma~\ref{lem:X}(\ref{item:X-min-V}), we have some set $\V\subseteq\reals^\Y_+$ and minimum representative sets $\R^* \subseteq \R$ and $\U^* \subseteq \U$, for $\ell$ and $L$ respectively, such that $\ell(\R^*) = \V = L(\U^*)$.
  %\jessiet{This is a different $\Sc$ than the first paragraph, right? The first $\Sc$ is representative for $\ell$.}\raft{They are both quantified separately, so not linked; are you saying it's confusing to use $\Sc$ here though because its role is slightly different?}\jessiet{Yes... attempting to clarify.}
  As $\R^*$ and $\U^*$ are miniumum, they cannot repeat loss vectors, and thus $|\R^*|=|\ell(\R^*)|$ and $|L(\U^*)|=|\U^*|$.
  We conclude that $\R^*$ and $\U^*$ are both in bijection with $\V$.
  The map $\varphi :\R^* \to \reals^d$, given by $\varphi(r) = u \in \U^*$ where $\ell(r) = L(u)$, is therefore well-defined.
  Condition (i) of an embedding is immediate.
  From Proposition~\ref{prop:representative-embeds-restriction}, $\ell$ embeds $\ell|_{\R^*}$ and $L$ embeds $L|_{\U^*}$, both via the identity embedding.
  Using condition (ii) from both embeddings, for all $p\in\simplex$ and $r\in\R^*$, we have
  \begin{equation*}
    r \in \gamma(p) \iff r \in \prop{\ell|_{\R^*}}(p) \iff \varphi(r) \in \prop{L|_{\U^*}}(p)
    \iff \varphi(r) \in \prop{L}(p)~,
  \end{equation*}
  giving condition (ii).
\end{proof}

We use this fact in the proof of Theorem~\ref{thm:discrete-loss-poly-embeddable} to show that every discrete loss is embedded by some polyhedral surrogate.
See Figure~\ref{fig:bayes-risks-01} for an illustration.

@@

In a strong sonse, the trim operation reduces a loss to its core: the unique minimal set of loss vectors that drive its statistical behavior.
One can therefore think of designing consistent convex surrogates as trying to ``fill out'' this minimal set with additional loss vectors so that one attains convexity while keeping trim the same.
(At least, the above is true when restricting to minimizable losses with polyhedral Bayes risks.)

@@

\section{Polyhedral Indirect Elicitation Implies Consistency}
\label{sec:poly-ie-consistency}
